Probabilistic Policy Violation Detection with IPFWD

### 2017-02-16 paper/main.tex
Refactor, remove IEEE notes, add bibliography, conclusion, related work

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

\author{Philip Nelson and Austin Voecks% <-this % stops a space
\thanks{Supported in part by Dell EMC Isilon through a grant and hardware}}

@@

\section{Introduction}

@@

This is a shift in mindset from typical firewalls. Instead of enforcing every
part of a security policy all the time, IPFWD enforces the policy some of the
time and extrapolates from the violations it encounters. This provides insight
into the violations that weren't detected. For this cost, you gain increased
firewall performance and network throughput in resource bound systems. It's
acceptable to allow a percentage of policy violations given that network
traffic patterns are often repeated and the goal is detection and not immediate
prevention. Preventative action may be taken later when resource requirements
are lower.

@@

% Firewall basics
\subsection{Firewall Basics}
Firewalls restrict the types of incoming and outgoing network traffic according
to a security policy. Without a security policy, administrators cannot know
what to allow and disallow. Typical firewalls enforce the policy by blocking
packets that do not match the rules. The default action for whitelist firewalls is to
block all traffic, additional rules describe exceptions. Blacklist firewalls
are the opposite in that by default they accept all traffic and additional
rules describe packet types that are not allowed. Since predicting and
blacklisting all possible attacks is infeasible, whitelist firewall rulesets
are the accepted standard and the focus of this paper.

% Why was this firewall chosen?
\subsection{Choice of IPFW}
IPFW is a stateful firewall written for FreeBSD which supports both IPv4 and
IPv6. It is comprised of several components: a kernel firewall filter rule
processor and its integrated packet accounting facility, the logging facility,
NAT, the dummynet traffic shaper, a forward facility, a bridge facility, and
an ipstealth facility.

IPFW was chosen for it's close integration with the FreeBSD operating system,
in kernel packet filter, and general performance. Performance is dependent on
system parameters and circumstances, but IPFW outperforms PF in stateful packet
filtering. Stateful rules are more powerful and flexible than stateless rules,
were no session information is maintained. Additionally, the capabilities of
IPFW extend beyond packet filtering into source based routing, traffic shaping
and more.

@@

% How does it work?
% - Probabilistic packet matching
Since IPFW uses a first-match-wins rule system, this early rule will be
encountered before the majority of the other rules in the rule set.  The
probability given in this rule determines the chance that we skip processing
all further rules, supposing that there hasn't been a match already.

@@

In a typical data center environment the vast majority of network traffic is
valid and going to be accepted eventually. However, these rule sets are often
complex and take increasing time to match a given packet depending on the
length. For a rule set of length $n$, we can expect to a packet to encounter
$\frac{n}{2}$ failed matches before matching the correct rule and being accepted.

@@

% - Log extension
While this method increases the speed the system processes packets, it
creates a chance that packets that would have normally been rejected will get
through the firewall. We'll refer to these types of packets as invald for the purpose of discussion.

@@

% When should this system be used?
\subsection{Application}
IPFWD is not intended to be used on all types of systems. It's application
instead most directly benefits the following types of systems:

\begin{enumerate}
    \item \textbf{High Network Performance}\\
      The gains provided by IPFWD are only apparent in systems were network
      throughput is matched by or out performs CPU speed. Systems that can
      already easily handle fully processing each packet will not benefit by
      reducing processing time. Given modern CPU clock speeds, systems with
      network cards slower than 1 Gb/sec will likely not benefit either.

    \item \textbf{Resource Conscious}\\
      In line with high performance systems, systems with strict resource
      restrictions can benefit by the reduced network related CPU usage.

    \item \textbf{Complex Firewall Requirements}\\
      IPFWD works by reducing the average number of rules that fail to match
      before a packet is finished being processed. Very simple rulesets will
      gain little from reducing the number of rules applied to each packet.

    \item \textbf{Detection Based Security Policy}\\ 
      By using IPFWD, administrators sacrifice the assurance that every invalid
      packet will be rejected.  However, they are still provided information as
      to whether some invalid packets were rejected, and that information can
      be used to infer the existance of additional invalid packets.

@@

\subsection{Rationale}
% Explain rationale for it acting as it does
IPFWD is based on the assumptions that the majority of network traffic is
valid, malicious events are rare, and TODO

@@

% Talk about security policy enforcement

% What does it mean for security if you're ignoring some of the packet?

% How do you trust this system?

% Talk about shift in mindset from "blocking attacks" to "detecting attacks"
% - Why is this reasonable?
% - Is this really different than traditional firewalls?

\section{Benchmarking Performance}

% Explain different systems
Tests were run on three platforms and five different network cards, resulting
in five different systems. The most important parameters to our tests are CPU,
NIC, and ENV. Descriptions of each system are given in the tables below:\\

@@

% WWU 10G
\begin{tabular}{ | l | L{19em} | } 
 \hline
 \multicolumn{2}{|c|}{ WWU 10Gb System } \\
 \hline
 \hline
 OS  &  FreeBSD 11 Release \\
 CPU &  8 x64 Intel\textsuperscript{\textregistered}\ Core\texttrademark\ i7-2600 CPU @ 3.4GHz \\
 MEM &  16Gb \\ 
 NIC &  ??? \\ 
 ENV &  Physical \\ 
 \hline
\end{tabular} \\\\

@@

% Virtualized 10G
\begin{tabular}{ | l | L{19em} | } 
 \hline
 \multicolumn{2}{|c|}{ Virtualized 10G System } \\
 \hline
 \hline
 OS  &  FreeBSD 11 Release \\
 CPU &  1 x64 Intel\textsuperscript{\textregistered}\ Xeon\textsuperscript{\textregistered}\ CPU @ 1.80GHz \\
 MEM &  512Mb \\
 NIC &  Ethernet 10Gbase-T \\ 
 ENV &  Hypervisor \\ 
 \hline
\end{tabular} \\\\

The systems tested all used x64 FreeBSD 11 Release, with varying CPU type,
CPU speed, RAM and network cards. Isilon and WWU systems were both physical
hardware, while the Virtualized system was in a hypervisor environment hosted
by a third party service.

Virtualization makes it more difficult to have the same confidence in the
results as with the physical machines. Tests on the Virtualized system were
repeated additional times to account for the variability of the underlying
physical hardware and contention between virtual machines.

The WWU and Virtualized tests were conducted on stock FreeBSD 11 without
additional network performance configuration applied. Isilon OneFS does contain
additional network performance tuning and demonstrates the additional gains
IPFWD can provide with careful FreeBSD 11 kernel configuration.

  \subsection{Netperf}
  % Explain tools
  Multiple network performance tools were explored but the final results were
  compiled using Netperf. Netperf was chosen because it was readily available
  on all test platforms, provides built-in CPU utilization measurement, and a
  wide range of test types.

  % Explain tests
  The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
  preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
  the target metric, which can be difficult to assess with TCP.

@@

  % Experimental method
  All tests were conducted using the following steps:
  \begin{enumerate}
      \item Machine 1
      \begin{enumerate}
          \item ipfw
          \item netserver
          \item CPU monitor
      \end{enumerate}
      \item Machine 2
      \begin{enumerate}
          \item ipfw
          \item ipfwd
          \item netperf
          \item CPU monitor
      \end{enumerate}
  \end{enumerate}

  All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules.
  Intuitively, larger production rule sets would benefit even further from
  IPFWD.

@@

  CPU utilization measurement is the most difficult parameter to measure in
  these tests. There are many factors that can account for variability between
  tests and jitter in the results. The Netperf manual describes these
  challenges and the variety of techniques used on different systems.

  Netperf was run in single threaded mode, which explains why throughput wasn't
  closer to the maximum throughput for the NIC on each system. The FreeBSD
  network stack allows multithreaded packet processing by utilizing multiple
  queues. Since we're indirectly testing the performance of IPFW through
  Netperf, there are two different CPU measurements to account for: kernel time
  and user time. Since IPFW runs in the kernel, it's CPU time is classified as
  kernel time.

  Netperf was able to measure CPU utilization on the Virtualized and WWU
  systems, but didn't work on the Isilon systems. Instead, ps was used on those
  systems.  These tests are comparable since both measured total system CPU
  utilization TODO

@@

  % throughput/CPU
  \subsection{Throughput / CPU}
  These graphs show network throughput divided by CPU utilization for each system. Higher values mean that the CPU time required to process packets is lower.

@@

  %  throughput & CPU
  \subsection{Throughput \& CPU}
  These graphs show the interaction between CPU utilization and network throughput for each system with respect to IPFWD's acceptance probability. The dashed lines are CPU utilization, solid is network throughput in Mb/second. Higher acceptance probability means a greater chance that a given packet is immediately accepted without further processing.

@@

  In CPU bound systems, IPFWD reduces the number of cycles it takes to process
  each packet, allowing more packets to be processed in the same time and
  increasing throughput. This is shown most clearly in the Virtualized system
  tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless of test
  type, the receiving system is always at 100\% CPU utilization. However, as
  the probability to accept increases in IPFWD the throughput also increases.
  This clearly shows that IPFWD decreases the CPU load required to process
  incoming traffic.

@@

@@

Firewall decision trees? (Complete Redundancy Detection)
TODO

% Machine learning firewalls

\section{Future Work}
% Firewall rule reordering so we can "fail as early as possible"
Ideally, a whitelist firewall's rule set would be constructed so that the most
commonly applied matches are reached earlier rather than later. IPFWD
accomplishes this task by artificial early stopping. Firewall rule reordering
is a could be another solution to this problem. Given logs of common network
traffic for the system, statistical analysis could show which rules are matched
more often than others. With this information, the rule set could be reordered
so that the most commonly applied rules are earliest.
TODO

\section{Conclusion}
TODO
% What performance benefit do we get?

% What security cost is there?

% Why is this useful?

% Why is this important?

\appendices
\section{Additional Figures}
TODO

% use section* for acknowledgement
\section*{Acknowledgment}
TODO

@@

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

>>>>

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}

\usepackage{blindtext}
\usepackage[style=ieee,backend=biber]{biblatex}
\usepackage[final]{microtype}
\usepackage{csquotes}
\usepackage[english]{babel} 
\usepackage{array}

\usepackage{hyperref}
\addbibresource{references.bib}

\usepackage{graphicx}
\graphicspath{ {images/} }

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\providecommand*{\opt}[1]{\texttt{#1}}
\providecommand*{\pkg}[1]{\textsf{#1}}

@@

\author{Philip Nelson and Austin Voecks% <-this % stops a space
\thanks{Supported in part by Dell EMC Isilon through grant and hardware}}

@@

% TODO rewrite
\section{Introduction}

@@

  This is a shift in mindset from typical firewalls \cite{networksecurity}.
  Instead of enforcing every part of a security policy all the time, IPFWD
  enforces the policy some of the time and extrapolates from the violations it
  encounters.  This provides insight into the violations that weren't detected.
  For this cost, you gain increased firewall performance and network throughput
  in resource bound systems. It's acceptable to allow a percentage of policy
  violations given that network traffic patterns are often repeated and the
  goal is detection and not immediate prevention. Preventative action may be
  taken later when resource requirements are lower.

@@

  % Firewall basics
  \subsection{Firewall Basics}

    Firewalls restrict the types of incoming and outgoing network traffic
    according to a security policy. Without a security policy, administrators
    cannot know what to allow and disallow. Typical firewalls enforce the
    policy by blocking packets that do not match the rules. The default action
    for whitelist firewalls is to block all traffic, additional rules describe
    exceptions. Blacklist firewalls are the opposite in that by default they
    accept all traffic and additional rules describe packet types that are not
    allowed. Since predicting and blacklisting all possible attacks is
    infeasible, whitelist firewall rulesets are the accepted standard and the
    focus of this paper.

  % Why was this firewall chosen?
  \subsection{Choice of IPFW}

    IPFW is a stateful firewall written for FreeBSD which supports both IPv4
    and IPv6. It is comprised of several components: a kernel firewall filter
    rule processor and its integrated packet accounting facility, the logging
    facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
    facility, and an ipstealth facility \cite{freebsdhandbook}.

    IPFW was chosen for it's close integration with the FreeBSD operating
    system, in kernel packet filter, and general performance. Performance is
    dependent on system parameters and circumstances, but IPFW outperforms PF
    in stateful packet filtering \cite{ipfwvspf}. Stateful rules are more
    powerful and flexible than stateless rules, were no session information is
    maintained \cite{networksecurity}.  Additionally, the capabilities of IPFW
    extend beyond packet filtering into source based routing, traffic shaping
    and more.

@@

  % How does it work?
  % - Probabilistic packet matching
  IPFWD's behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them \cite{TODO}.

  Since IPFW uses a first-match-wins rule system, this early rule will be
  encountered before the majority of the other rules in the rule set.  The
  probability given in this rule determines the chance that we skip processing
  all further rules, supposing that there hasn't been a match already.

@@

  In a typical data center environment the vast majority of network traffic is
  valid and going to be accepted eventually \cite{TODO}. However, these rule
  sets are often complex and take increasing time to match a given packet
  depending on the length. For a rule set of length $n$, we can expect to a
  packet to encounter $\frac{n}{2}$ failed matches before matching the correct
  rule and being accepted.

@@

  % - Log extension
  While this method increases the speed the system processes packets, it
  creates a chance that packets that would have normally been rejected will get
  through the firewall. We'll refer to these types of packets as invalid for
  the purpose of discussion.

@@

  % When should this system be used?
  \subsection{Application}

    IPFWD is not intended to be used on all types of systems. It's application
    instead most directly benefits the following types of systems:

    \begin{enumerate}
      \item \textbf{High Network Performance}

        The gains provided by IPFWD are only apparent in systems were network
        throughput is matched by or out performs CPU speed. Systems that can
        already easily handle fully processing each packet will not benefit by
        reducing processing time. Given modern CPU clock speeds, systems with
        network cards slower than 1 Gb/sec will likely not benefit either.

      \item \textbf{Resource Conscious}

        In line with high performance systems, systems with strict resource
        restrictions can benefit by the reduced network related CPU usage.

      \item \textbf{Complex Firewall Requirements}

        IPFWD works by reducing the average number of rules that fail to match
        before a packet is finished being processed. Very simple rulesets will
        gain little from reducing the number of rules applied to each packet.

      \item \textbf{Detection Based Security Policy}

        By using IPFWD, administrators sacrifice the assurance that every
        invalid packet will be rejected.  However, they are still provided
        information as to whether some invalid packets were rejected, and that
        information can be used to infer the existence of additional invalid
        packets.

@@

@@

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization
  \cite{TODO}. At the network and host level, this is implemented in part by a
  firewall rule set. A policy violation on an internal network is more serious
  than an external network. Internet facing hosts can expect to be scanned,
  experience network anomalies and be attacked by third party agents
  semi-regularly.

  However when this happens on an internal network, there is a strong
  possibility something else is going wrong. The usage of IPFWD is predicated
  on the assumption that it's sufficient to be notified of these events so
  further action can be taken later than to always immediately stop them.

  % What does it mean for security if you're ignoring some of the packet?
  Allowing some percentage of network packets through the firewall without
  inspection does mean that attacks that would have normally been blocked could
  succeed. In addition to performance improvements, the following factors help
  to account for this risk:

  \begin{enumerate}

    % Bad things are often repeated
    \item \textbf{Repetition}

      Malicious network activity and network errors are likely to be repeated
      over time \cite{TODO}. The longer the activity persists, the more likely
      it is to be detected by IPFW even when IPFWD is allowing a high
      percentage of the network traffic through unchecked.

    \item \textbf{Disruption}

      Though some malicious network activity may only consist of a single
      packet, others will require multiple packets. Since IPFW works on a
      per-packet basis, malicious network activity will experience
      approximately $m\%$ packet loss where $m$ is IPFWD's probability of early
      acceptance. The exact effect this would have on malware or reconnaissance
      would depend, but generally will increase the chance they fail or produce
      invalid results.

  % How do you trust this system?
  Knowledge of a network's traffic patterns and the additional log information
  provided by IPFWD can help. It allows administrators to extrapolate the
  existence of undetected network events through simple statistical models.

  % Talk about shift in mindset from "blocking attacks" to "detecting attacks"
  % - Why is this reasonable?
  % - Is this really different than traditional firewalls?
  TODO

  % example application
  \subsection{Example Scenario}

    Suppose each host in a data center's internal network is running IPFWD on
    each host, average CPU usage is $75\%$, and have rule sets blocking common
    network attacks. One of the hosts has been compromised and begins
    conducting network reconnaissance by port scanning the other hosts on the
    network.

    Firewall rules are in place to block host to host communication $m\%$ of
    the ports being scanned. This means that $1 - m\%$ of the scan would be
    allowed regardless of IPFWD's action. This percentage represents valid
    host-to-host communication as determined by the security policy.

    IPFWD would allow $75\%$ of the policy violating port scan through to the
    host, and $75\%$ of the replies back to the compromised host. This results
    in $75\% * 75\% = 56.25\%$ of the port scan that should have been stopped
    succeeding.  However, the other $1 - 56.25 = 43.75\%$ of the port scan was
    blocked and logged, alerting system administrators to the presence of
    malicious network activity so further action can be taken.

  % describe difference between reply and no reply violations
  \subsection{Security Cost}

    The chance of missing a security policy violation depends on the type of
    traffic. The following situations exist when both incoming and outgoing
    firewall rules are employed:

    \begin{enumerate}

      % only one chance to catch
      \item \textbf{Incoming or Outgoing + No Reply}

        Here, the firewall only has one chance to catch the policy violation.
        So if the early acceptance probability is $m$, then $m\%$ of the policy
        violating packets will be allowed.

      % two chances to catch 
      \item \textbf{Incoming or Outgoing + Reply}

        Here, the firewall has two chances to catch the policy violation.  So
        if the early acceptance probability is $m$, then $m'\%$ of the policy
        violating packets will be allowed where $m' = m * m$.

% Explain different systems and tests
\section{Benchmarking Performance}

  Tests were run on three platforms and five different network cards, resulting
  in five different systems. The most important parameters to our tests are
  CPU, NIC, and ENV. Descriptions of each system are given in the tables
  below:\\

@@

@@

  % Virtualised 10G
  \begin{tabular}{ | l | L{19em} | } 
    \hline
    \multicolumn{2}{|c|}{ Virtualised 10G System } \\
    \hline
    \hline
    OS  &  FreeBSD 11 Release \\
    CPU &  1 x64 Intel\textsuperscript{\textregistered}\ Xeon\textsuperscript{\textregistered}\ CPU @ 1.80GHz \\
    MEM &  512Mb \\
    NIC &  Ethernet 10Gbase-T \\ 
    ENV &  Hypervisor \\ 
    \hline
  \end{tabular} \\\\

  The systems were tested on x64 FreeBSD 11 Release, with varying CPU type, CPU
  speed, RAM and network cards. Isilon and WWU systems were both physical
  hardware, while the Virtualised system was in a hypervisor environment hosted
  by a third party service.

  Virtualisation makes it more difficult to have the same confidence in the
  results as with the physical machines. Tests on the Virtualised system were
  repeated additional times to account for the variability of the underlying
  physical hardware and contention between virtual machines.

  The WWU and Virtualised tests were conducted on stock FreeBSD 11 without
  additional network performance configuration applied. Isilon OneFS does
  contain additional network performance tuning and demonstrates the additional
  gains IPFWD can provide with careful FreeBSD 11 kernel configuration.

  \subsection{Netperf}

    % Explain tools
    Multiple network performance tools were explored but the final results were
    compiled using Netperf. Netperf was chosen because it was readily available
    on all test platforms, provides built-in CPU utilization measurement, and a
    wide range of test types \cite{netperf}.

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
    the target metric, which can be difficult to assess with TCP
    \cite{TODO}.

@@

    % Experimental method
    All tests were conducted using the following steps:
    \begin{enumerate}
        \item Machine 1
        \begin{itemize}
          \item \texttt{ipfw}
          \item \texttt{netserver}
          \item \texttt{CPU monitor}
        \end{itemize}
        \item Machine 2
        \begin{itemize}
          \item \texttt{ipfw}
          \item \texttt{ipfwd}
          \item \texttt{netperf}
          \item \texttt{CPU monitor}
        \end{itemize}
    \end{enumerate}

    All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules
    depending on the environment and network card configuration. Larger
    production rule sets would benefit even further from IPFWD.

@@

    CPU utilization measurement is the most difficult parameter to measure in
    these tests. There are many factors that can account for variability
    between tests and jitter in the results. The Netperf manual describes these
    challenges and the variety of techniques used on different systems.
    \cite{netperf}

    Netperf was run in single threaded mode, which explains why throughput
    wasn't closer to the maximum throughput for the NIC on each system. The
    FreeBSD network stack allows multithreaded packet processing by utilizing
    multiple queues \cite{freebsdhandbook}. Since we're indirectly testing the
    performance of IPFW through Netperf, there are two different CPU
    measurements to account for: kernel time and user time. Since IPFW runs in
    the kernel, it's CPU time is classified as kernel time.

    Netperf was able to measure CPU utilization on the Virtualized and WWU
    systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
    used. These tests are comparable since both measured user CPU utilization,
    not kernel CPU utilization.

    In systems where the throughput of the NIC exceeds what the CPU can supply,
    CPU utilization is high and throughput is lower than the NIC maximum (eg:
    Virtualised system). Conversely, when CPU availablility exceeds NIC
    throughput, throughput is high and CPU utilization is low (eg: WWU 1G
    system).

@@

  %  throughput & CPU
  \subsection{Throughput \& CPU}

    These graphs show the interaction between CPU utilization and network
    throughput for each system with respect to IPFWD's early acceptance
    probability. The dashed lines are CPU utilization, solid is network
    throughput in Mb/second. Higher acceptance probability means a greater
    chance that a given packet is immediately accepted without further
    processing.

    The upward trend in throughput on the x-axis shows that as the early
    acceptance probability increases, so does throughput. The CPU usage shows
    that this throughput is gained without additional CPU utilization, meaning
    the system is spending fewer cycles to process each packet on average.

    The lack of any trend in the WWU 1G graph is expected and addressed in the
    \textit{Discussion} section.

  % throughput/CPU
  \subsection{Throughput / CPU}

    These graphs show network the ratio of throughput to CPU utilization for
    each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
    presented here is the same as the previous graphs, shown differently. The
    upward trend over the x-axis shows that as the probability of early
    acceptance increases, the system is spending less CPU time to process
    packets on average.

    The dashed line shows the line of best fit for the throughput to CPU
    utilization ratio and shows more clearly the upward trend.

@@

@@

    In CPU bound systems, IPFWD reduces the number of cycles it takes to
    process each packet, allowing more packets to be processed in the same time
    and increasing throughput. This is shown most clearly in the Virtualised
    system tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless
    of test type, the receiving system is always at 100\% CPU utilization.
    However, as the probability to accept increases in IPFWD the throughput
    also increases.  This clearly shows that IPFWD decreases the CPU load
    required to process incoming traffic.

@@

    % WWU 1G
    However, the WWU 1G tests provide an example of a NIC bound system. The
    lack of any performance gains in the WWU 1G tests provides an example of
    when IPFWD should not be used. In this system, the CPU and other system
    resources are more than sufficient to apply the entire IPFW ruleset to each
    packet at wire-speed. Reducing the amount of CPU time it takes to process
    each packet does not provide any benefit.

@@

  % Similiar work
  Related work \cite{exploitingpacketsampling, analysisnetflow,
  monitoringpacketsampling} into packet sampling in the space of network
  security has been focused on detecting and classifying threats, not
  performance enhancement.

  %Firewall decision trees? (Complete Redundancy Detection)

\section{Future Work}

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a whitelist firewall's rule set would be constructed so that the
  most commonly applied matches are reached earlier rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  is a could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest.  TODO

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets

  % have detection of bad traffic reduce acceptance rate

\section{Conclusion}
  % What performance benefit do we get?
  % What security cost is there?
  % Why is this useful?
  % Why is this important?

  In performance sensitive computing environments, complex firewall rules
  combined with high network throughput can result in heavy CPU load. IPFWD
  addresses this problem by providing a graceful tradeoff between packet
  visibility, total security policy enforcement and system performance.

  IPFWD reduces the average amount of work to process packets by adding an
  early rule to IPFD that has a chance to accept any packet. The security
  implications of this behavior have been addressed and the performance
  benefits provided have been demonstrated in a variety of systems.

  IPFWD can be directly applied to high network performance and resource
  conscious systems with complex firewall requirements to reduce CPU load and
  increase network throughput.

  This work also serves as a proof of concept for future probabilistic packet
  matching firewall implementations.

\appendices
\section{Additional Information}
  TODO link to github repository

% use section* for acknowledgement
%\section*{Acknowledgment}
% TODO

@@

% references section
\printbibliography
